---
output:
  html_document:
    self_contained: no
---

Digital On-Boarding 
-------------------


_**6th report iteration**_

```{r, echo = FALSE}
# This is the first mandatory section.

title     <- "[Digital On-boarding] : What channel is used for engagement of products? Customer profile: culture, volunteering and traveling."

# Keywords. Text separated by commas. E.g. keywords <- 'customer health, recibos' 
# Keywords can classify the attribute in categories such as 'valor, vinculación, digitalidad', describe used tables such as 'recibos, 
# transacciones tarjetas, segmento plan uno' or include other relevant info. Search keywords in the portal to find commonly used terms
# and create new keywords only if justified.

keywords  <- 'digitalidad, digitalization, contratacion tarjetas, contracting cards, contratacion cuenta corriente, contracting current account, bbva.es, nuevo cliente, new client, alta cliente, client registration, perfil cliente, customer profile, voluntariado, volunteering, cultura, culture, traveling'  
```


```{r, echo=FALSE}
# This is the second mandatory section.

suppressMessages(library(DBI))    # This avoids loading messages and warnings showing up
suppressMessages(library(rJava))
options(warn=-1)
source('~/bda_clarity/tools/warehouse_basics.R')
```


```{r, echo=FALSE}
# This is the second mandatory section.

suppressMessages(library(DBI))    # This avoids loading messages and warnings showing up
suppressMessages(library(rJava))
suppressMessages(library(ggplot2))
suppressMessages(library(grid))
suppressMessages(library(hexbin))
suppressMessages(library(lattice))
suppressMessages(library(knitr))
suppressMessages(library(googleVis))
suppressMessages(library(data.table))
suppressMessages(library(plyr))
op <- options(gvis.plot.tag="chart")
`%ni%` <- Negate(`%in%`)
options(warn=-1, scipen=3, width=450)
source('~/bda_clarity/tools/warehouse_basics.R')

```

``` {r echo=FALSE}

multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  require(grid)
  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
} ; 




chname <- function(df)
{
  rex <- '^[[:alnum:]_]+\\.([[:alnum:]_]+)$'
  nam <- colnames(df)
  ix  <- which(grepl(rex, nam))
  
  nam[ix] <- gsub(rex, '\\1', nam[ix])
  
  colnames(df) <- nam
  
  df
};

```

### PREFERENCES OF CLIENTS 

```{r echo=FALSE, cache=TRUE, eval=FALSE}

do.hive(" CREATE TABLE IF NOT EXISTS elenak.digital_onboard_culture AS
         
           SELECT aa.cod_persona,   
                  bb.ultimo_mes,
                  bb.ultimos_3_meses,
                  bb.ultimos_6_meses,
                  bb.ultimos_12_meses
         
           FROM elenak.altas_digitales_2014 aa 
         
           JOIN clarity_elements.num_operaciones_cultura bb ON aa.cod_persona = bb.cod_persona  "); 


do.hive(" CREATE TABLE IF NOT EXISTS elenak.digital_onboard_volunteering AS
         
           SELECT aa.cod_persona,   
                  bb.ultimo_mes,
                  bb.ultimos_3_meses,
                  bb.ultimos_6_meses,
                  bb.ultimos_12_meses
         
           FROM elenak.altas_digitales_2014 aa 
         
           JOIN clarity_elements.num_operaciones_volunt_recib bb ON aa.cod_persona = bb.cod_persona  "); 


do.hive(" CREATE TABLE IF NOT EXISTS elenak.digital_onboard_traveling AS
         
           SELECT aa.cod_persona,   
                  bb.ultimo_mes,
                  bb.ultimos_3_meses,
                  bb.ultimos_6_meses
         
           FROM elenak.altas_digitales_2014 aa 
         
           JOIN clarity_elements.num_operaciones_agencias_viajes bb ON aa.cod_persona = bb.cod_persona  "); 

```

Para 2011 clientes con altas digitales en 2014 hemos visto cúantos de ellos pueden referirse a tales sectores de gustos como Cultura, Voluntariado y Travalers.

**Cultura**: según la actividad del cliente durante el último mes entran en la bolsa **`r qhive("select count(distinct cod_persona) from elenak.digital_onboard_culture where ultimo_mes <> 0")`** personas y según la actividad durante los últimos tres meses entran **`r qhive("select count(distinct cod_persona) from elenak.digital_onboard_culture where ultimos_3_meses <> 0")`** personas.

**Voluntariado**: según la actividad del cliente durante el último mes entran en la bolsa **`r qhive("select count(distinct cod_persona) from elenak.digital_onboard_volunteering where ultimo_mes <> 0")`** personas y según la actividad durante los últimos tres meses entran **`r qhive("select count(distinct cod_persona) from elenak.digital_onboard_volunteering where ultimos_3_meses <> 0")`** personas.

**Travalers**: según la actividad del cliente durante el último mes entran en la bolsa **`r qhive("select count(distinct cod_persona) from elenak.digital_onboard_traveling where ultimo_mes <> 0")`** personas y según la actividad durante los últimos tres meses entran **`r qhive("select count(distinct cod_persona) from elenak.digital_onboard_traveling where ultimos_3_meses <> 0")`** personas.

De este modo al ver el plazo de un mes acoplaríamos para Cultura 3,43% de clientes, para Voluntariado sólo 0,45% y para Travelers 0,05%. 




