---
title: "[PYMES] ComEx exploration"
author: "Maria Hernandez"
date: "8/5/2015"
output: 
  html_document:
    self_contained: no
---

```{r, echo=FALSE}
# This is the first mandatory section.

title     <- '[Pymes]: Exploration of ComEx activity'
keywords  <- 'pymes,comex,comercio,exterior'
```

```{r mandatory, echo=FALSE}
suppressMessages(library(DBI))    # This avoids loading messages and warnings showing up
suppressMessages(library(rJava))
options(warn=-1)

source('~/bda_clarity/tools/warehouse_basics.R')
source('~/bda_clarity/tools/methods_connect.R')
source('~/bda_clarity/tools/write.hive.R')
```

```{r libraries, echo=FALSE}
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(plyr))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(reshape))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(googleVis))
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(reshape))
```

```{r environment}
month <- '20151130'
```

```{r dependencies, echo=FALSE}
contrapartida <- clarity.use_table(DEPENDENCY_LASTPART_CATALOGO, 'contrapartida')

global <- clarity.use_table(DEPENDENCY_LASTPART_CATALOGO, 'segmentacion_global', 
                            select='cod_sgmento, des_sgmento')

segmento <- clarity.use_table(DEPENDENCY_LASTPART_PRO_TABLE, 'segmento_global')

intervinientes <- clarity.use_table(DEPENDENCY_LASTPART_PRO_TABLE, 'intervinientes_corp')

clientes <- clarity.use_table(DEPENDENCY_LASTPART_PRO_TABLE, 'clientes_corp')

movimientos <- clarity.use_table(DEPENDENCY_FULL_PRO_TABLE, 'movimientos_cuentas_corp')
```


*Objective:* In this report we are building the HasComEx indicator inside BBVA. 

### How can we see ComEx *inside* the bank?

There are several FX products and different ways they are found in the data:

* Avales y garnatías internacionales.
    + Epígrafe de gestión 3315321
* Créditos Documentarios. 
    + Contrapartida 0239: CREDITOS DOCUMENTARIOS DE EXPORTACION o epígrafe de gestión 3355012
    + Contrapartida 0243: CREDITOS DOCUMENTARIOS DE IMPORTACION o epígrafe de gestión 3355017
    + Contrapartida 0244: LINEAS DE CREDITOS DOCUMENTARIOS
* ECA - Financiación Comercio Exterior (FCI): 
    + Contrapartida 0247: FCI  (REFIS) (NO) VENCIDOS, unidad 2130. 
* Medios de pagos internacionales
    + Cheques Internacionales Emitidos: epígrafe de gestión 3356001
    + Cheques Internacionales Recibidos: epígrafe de gestión 3356002 
    + Órdenes de pago Emitidas: epígrafe de gestión 3356101 
    + Órdenes de pago Recibidas: epígrafe de gestión 3356130 
* Remesas
    + Remesas de Importación: epígrafe de gestión 3355010, 3355011
    + Remesas de Importación: epígrafe de gestión 3355015, 3355016
    + Contrapartida 0242: OPERACIONES DOCUMENTARIAS COMEX (remesas, aplicativos prisma comex)
    + Otras Remesas y Créditos documentarios: epígrafe de gestión 3355698
* Trade Loan
    + Anticipo import/export: 
        - ¿? epígrafe de gestión 3305410
        - Import: 3305811
        - Export: 3305816 
    + Forfaiting. Contrapartida 0248: FORFAITING CONTRATOS VENCIDO O NO
    + Financiación divisa. Contrapartidas  
        - 0250: FINANCIACION EN DIVISA - PROPUESTA O L?M
        - 0251: FINANCIACION EN DIVISA - OPERACIONES
        - 0252: FINANCIACION EN DIVISA - OPERACIONES
        - Exportación: epígrafe de gestión 3306731
        - Importación: epígrafe de gestión 3306726


```{r createContrapartidaComex, eval=FALSE}
do.hive("drop table da_pymes.contrapartida_comex")

q <- paste0("create table da_pymes.contrapartida_comex as 
  select cod_pgccontr as cod_contrapartida, max(des_ctrpartd) as des_contrapartida
  from da_catalogos.contrapartida 
  where cod_pgccontr in 
    ('0239', '0243', '0244', '0247', '0242', '0250', '0251', '0252', '0248')
      and partition_id = '", month, "'
  group by cod_pgccontr
")
do.hive(q)
do.impala("invalidate metadata da_pymes.contrapartida_comex")

```

```{r getContrapartidaComex}
q <- "select * from da_pymes.contrapartida_comex"
dd <- qimpala(q)
```


```{r createProdGestComex, eval=FALSE}

partida <- c(15321, 
             56001, 56030, 
             56101, 56130, 
             55010, 55011, 55015, 55016, 
             55012, 55017, 55698,
             05410, 05811, 05816)
estado <- rep(33, length(partida))
descripcion <- c("Avales y Garantías Internacionales",
                 "Cheques Internacionales Emitidos",
                 "Cheques Internacionales Recibidos", 
                 "Órdenes de pago Emitidas", 
                 "Órdenes de pago Recibidas",
                 "Remesas de Importación",
                 "Remesas de Importación", 
                 "Remesas de Exportación", 
                 "Remesas de Exportación",
                 "Créditos documentarios de Importación",
                 "Créditos documentarios de Exportación",
                 "Otras remesas y créditos documentarios",
                 "Anticipos Import/Export", 
                 "Anticipos Import", 
                 "Anticipos Export"
                 )

epigrafe_comex <- data.frame(estado, partida, descripcion)
do.hive("drop table da_pymes.epigrafe_comex")

write.hive(epigrafe_comex, "da_pymes.epigrafe_comex")
do.impala("invalidate metadata da_pymes.epigrafe_comex")

do.hive("drop table da_pymes.prod_gest_comex")

q <- paste0("create table da_pymes.prod_gest_comex as 
  select cast(cod_prodct as int) as cod_prodct, estado as cod_std, partida as cod_prtda, descripcion
  from da_pro.catalogo_productos_gestion_epigrafes_relacion 
    join da_pymes.epigrafe_comex ep on ep.estado = cast(trim(cod_std) as int) and ep.partida = cast(trim(cod_prtda) as int) 
    and partition_id = '", month, "'
")
do.hive(q)
do.impala("invalidate metadata da_pymes.prod_gest_comex")

```

```{r getProdGestComex}

q <- "select * from da_pymes.prod_gest_comex"
dd <- qimpala(q)
```


## Quick Exploration

```{r segmentoGlobalCatalogo, cache=TRUE, echo=FALSE}
seg <- qimpala("select cast(trim(cod_sgmento) as int) cod_global,
               max(des_sgmento) as des_global
from da_catalogos.segmentacion_global 
where partition_id = '", month, "'
group by cast(trim(cod_sgmento) as int)
")
```

### Contrapartidas

Let's see the volumen of this type of contracts.

```{r comexClients, eval=FALSE}
do.hive("drop table da_pymes.clientes_comex_contrapartidas")
q <- paste0("create table da_pymes.clientes_comex_contrapartidas as
select cod_entalfa, cod_persctpn, cod_idcontra, cod_pgccontr, des_contrapartida
from da_pymes.contrapartida_comex as contr 
  join da_pro.intervinientes_corp ic on trim(cod_pgccontr)= cod_contrapartida
where partition_id= '", month, "' and trim(cod_ctippe)='TIT'
")
do.hive(q)
do.impala("invalidate metadata da_pymes.clientes_comex_contrapartidas")

```

We obtain how many clients and contracts there are for each contrapartida, depending on the customer being Legal or Natural and their Global Segment.

```{r numeros, echo=FALSE, eval=FALSE}
# 84.274 - 84.274
qimpala(paste0("select cod_contrapartida, count(*), count(distinct ic.cod_idcontra)
from da_pymes.contrapartida_comex cc
   join da_pro.productos_contratados ic on trim(cc.cod_contrapartida) = trim(ic.cod_pgccontr) and ic.partition_id = '", month, "'  and ic.cod_entalfa = '0182'
group by cod_contrapartida order by cod_contrapartida
  "))

# 38 -26
qimpala("select count(*), count(distinct ic.cod_idcontra)
 from da_pymes.contrapartida_comex cc
   join da_pro.productos_contratados_detalle ic on trim(cc.cod_contrapartida) = trim(ic.cod_pgccontr) and ic.partition_id = '20150630' and ic.cod_entalfa = '0182'
  ")

#1    48989                           38976
qimpala("select count(*), count(distinct ic.cod_idcontra)
from da_pymes.contrapartida_comex cc
   join da_pro.intervinientes_corp ic on trim(cc.cod_contrapartida) = trim(ic.cod_pgccontr) and ic.partition_id = '20150630' and ic.cod_entalfa = '0182'
  join da_pro.productos_contratados pc on trim(ic.cod_idcontra) = trim(pc.cod_idcontra) and pc.partition_id = '20150630' and pc.cod_entalfa = '0182'
  ")

#1     8067                            4066
qimpala("select count(*), count(distinct ic.cod_idcontra)
from da_pymes.contrapartida_comex cc
   join da_pro.intervinientes_corp ic on trim(cc.cod_contrapartida) = trim(ic.cod_pgccontr) and ic.partition_id = '20150630' and ic.cod_entalfa = '0182'
  join da_pro.productos_contratados_detalle pc on trim(ic.cod_idcontra) = trim(pc.cod_idcontra) and pc.partition_id = '20150630' and pc.cod_entalfa = '0182'
  ")

# Cruce al revés: filtro por contrapartida en pc y join con intervinientes
#1    48989                           38976
# lo mismo que antes
qimpala("select count(*), count(distinct ic.cod_idcontra)
from da_pymes.contrapartida_comex cc
   join da_pro.productos_contratados ic on trim(cc.cod_contrapartida) = trim(ic.cod_pgccontr) and ic.partition_id = '20150630' and ic.cod_entalfa = '0182'
  join da_pro.intervinientes_corp pc on trim(ic.cod_idcontra) = trim(pc.cod_idcontra) and pc.partition_id = '20150630' and pc.cod_entalfa = '0182'
  ")

# Cruce al revés: filtro por contrapartida en pcd y join con intervinientes
# 0 - 0
qimpala("select count(*), count(distinct ic.cod_idcontra)
from da_pymes.contrapartida_comex cc
   join da_pro.productos_contratados_detalle ic on trim(cc.cod_contrapartida) = trim(ic.cod_pgccontr) and ic.partition_id = '20150630' and ic.cod_entalfa = '0182'
  join da_pro.intervinientes_corp pc on trim(ic.cod_idcontra) = trim(pc.cod_idcontra) and pc.partition_id = '20150630' and pc.cod_entalfa = '0182'
  ")

# contrapartidas de pc de los contratos de intervinientes: 
#   cod_pgccontr count(*) count(distinct ic.cod_idcontra)
# 1         0248     4240                            4157
# 2         0239     3788                            3788
# 3         0247    14460                            4530
# 4         0242     7970                            7970
# 5         0252     5262                            5262
# 6         0250      819                             819
# 7         0251      805                             805
# 8         0243    11645                           11645
qimpala("select pc.cod_pgccontr, count(*), count(distinct ic.cod_idcontra)
from da_pymes.contrapartida_comex cc
   join da_pro.intervinientes_corp ic on trim(cc.cod_contrapartida) = trim(ic.cod_pgccontr) and ic.partition_id = '20150630' and ic.cod_entalfa = '0182'
  join da_pro.productos_contratados pc on trim(ic.cod_idcontra) = trim(pc.cod_idcontra) and pc.partition_id = '20150630' and pc.cod_entalfa = '0182'
group by pc.cod_pgccontr
  ")

# contrapartidas de pcd de los contratos de intervinientes: todas NULL
qimpala("select pc.cod_pgccontr, count(*), count(distinct ic.cod_idcontra)
from da_pymes.contrapartida_comex cc
   join da_pro.intervinientes_corp ic on trim(cc.cod_contrapartida) = trim(ic.cod_pgccontr) and ic.partition_id = '20150630' and ic.cod_entalfa = '0182'
  join da_pro.productos_contratados_detalle pc on trim(ic.cod_idcontra) = trim(pc.cod_idcontra) and pc.partition_id = '20150630' and pc.cod_entalfa = '0182'
group by pc.cod_pgccontr
  ")

# contrapartidas de ic de los contratos de pc: 
#   cod_pgccontr count(*) count(distinct ic.cod_idcontra)
# 1         0248     4240                            4157
# 2         0239     3788                            3788
# 3         0247    14460                            4530
# 4         0242     7970                            7970
# 5         0252     5262                            5262
# 6         0250      819                             819
# 7         0251      805                             805
# 8         0243    11645                           11645
qimpala("select pc.cod_pgccontr, count(*), count(distinct ic.cod_idcontra)
from da_pymes.contrapartida_comex cc
   join da_pro.productos_contratados ic on trim(cc.cod_contrapartida) = trim(ic.cod_pgccontr) and ic.partition_id = '20150630' and ic.cod_entalfa = '0182'
  join da_pro.intervinientes_corp pc on trim(ic.cod_idcontra) = trim(pc.cod_idcontra) and pc.partition_id = '20150630' and pc.cod_entalfa = '0182'
group by pc.cod_pgccontr
  ")


# contrapartidas de ic de los contratos de pcd: 
qimpala("select ic.cod_pgccontr, count(*), count(distinct ic.cod_idcontra)
from da_pymes.contrapartida_comex cc
   join da_pro.productos_contratados_detalle pc on trim(cc.cod_contrapartida) = trim(pc.cod_pgccontr) and pc.partition_id = '20150630' and pc.cod_entalfa = '0182'
  join da_pro.intervinientes_corp ic on trim(ic.cod_idcontra) = trim(pc.cod_idcontra) and ic.partition_id = '20150630' and ic.cod_entalfa = '0182'
group by ic.cod_pgccontr
  ")

# cómo son los contratos que se pierden al cruzar prod.contratados con intervinientes por cod_idcontra
qimpala("select substring(cod_idcontra, 1, 6) inicio, idcontra2 is null as idcontra_null, count(*), count(distinct cod_idcontra)
        from (
   select pc.*, ic.cod_idcontra as idcontra2
 from da_pymes.contrapartida_comex cc
    join da_pro.productos_contratados pc on trim(cc.cod_contrapartida) = trim(pc.cod_pgccontr) and pc.partition_id = '20150630' and pc.cod_entalfa = '0182'
    left join da_pro.intervinientes_corp ic on trim(pc.cod_idcontra) = trim(ic.cod_idcontra) and ic.partition_id = '20150630' and ic.cod_entalfa = '0182'
   ) x 
 group by substring(cod_idcontra, 1, 6) , idcontra2 is null
")


# cómo son los contratos que se pierden al cruzar prod.contratados con intervinientes por cod_idcontra
#    cod_pgccontr idcontra2_is_null     n     s          p
#  1:         0239                 0  3788  4852 0.78070899
#  2:         0239                 1  1064  4852 0.21929101
#  3:         0242                 0  7970 24105 0.33063680
#  4:         0242                 1 16135 24105 0.66936320
#  5:         0243                 0 11645 29850 0.39011725
#  6:         0243                 1 18205 29850 0.60988275
#  7:         0244                 1 17495 17495 1.00000000
#  8:         0247                 0 14460 36409 0.39715455
#  9:         0247                 1 21949 36409 0.60284545
# 10:         0248                 0  4240 34772 0.12193719
# 11:         0248                 1 30532 34772 0.87806281
# 12:         0250                 0   819  9259 0.08845448
# 13:         0250                 1  8440  9259 0.91154552
# 14:         0251                 0   805 75500 0.01066225
# 15:         0251                 1 74695 75500 0.98933775
# 16:         0252                 0  5262 46279 0.11370168
# 17:         0252                 1 41017 46279 0.88629832

dd <- qimpala("select cod_pgccontr, idcontra2 is null, count(*)
        from (
   select ic.*, pc.cod_idcontra as idcontra2
 from da_pymes.contrapartida_comex cc
    join da_pro.intervinientes_corp ic on trim(cc.cod_contrapartida) = trim(ic.cod_pgccontr) and ic.partition_id = '20150630' and ic.cod_entalfa = '0182'
    left join da_pro.productos_contratados pc on trim(pc.cod_idcontra) = trim(ic.cod_idcontra) and pc.partition_id = '20150630' and ic.cod_entalfa = '0182'
   ) x 
group by cod_pgccontr, idcontra2 is null
")
dd2 <- data.table(dd)[,s:=sum(n),by=c("cod_pgccontr")]
dd2$p <- dd2$n/dd2$s

dd <- qimpala("select cod_pgccontr, idcontra2 is null, count(*)
        from (
   select ic.*, pc.cod_idcontra as idcontra2
 from da_pymes.contrapartida_comex cc
    join da_pro.intervinientes_corp ic on trim(cc.cod_contrapartida) = trim(ic.cod_pgccontr) and ic.partition_id = '20150630' and ic.cod_entalfa = '0182'
    left join da_pro.productos_contratados_detalle pc on trim(pc.cod_idcontra) = trim(ic.cod_idcontra) and pc.partition_id = '20150630' and ic.cod_entalfa = '0182'
   ) x 
group by cod_pgccontr, idcontra2 is null
")


```

```{r getContrapartidas}
query <- paste0("select cod_segmsubo, xti_persona, cod_pgccontr, cc.des_contrapartida, 
  count(*) ncontr, count(distinct cli.cod_persctpn) ncli
from da_pymes.clientes_comex_contrapartidas cli 
  join da_pymes.contrapartida_comex cc on cc.cod_contrapartida = cli.cod_pgccontr
  join da_pro.segmento_global sg on sg.cod_persctpn = cli.cod_persctpn and sg.partition_id = '", month, "' and sg.cod_entalfa = '0182'
  join da_pro.clientes_corp corp on corp.cod_persctpn = cli.cod_persctpn and corp.partition_id = '", month, "' and corp.cod_entalfa = '0182'
group by cod_segmsubo, xti_persona, 
  cod_pgccontr, cc.des_contrapartida
")
contr <- qimpala(query)
names(contr) <- c("global", "tipo_persona", "cod", "contrapartida", "ncontr", "ncli")

query <- paste0("select cod_segmsubo, xti_persona,
  count(*) ncontr, count(distinct cli.cod_persctpn) ncli
from da_pymes.clientes_comex_contrapartidas cli 
  join da_pymes.contrapartida_comex cc on cc.cod_contrapartida = cli.cod_pgccontr
  join da_pro.segmento_global sg on sg.cod_persctpn = cli.cod_persctpn and sg.partition_id = '", month, "' and sg.cod_entalfa = '0182'
  join da_pro.clientes_corp corp on corp.cod_persctpn = cli.cod_persctpn and corp.partition_id = '", month, "' and corp.cod_entalfa = '0182'
group by cod_segmsubo, xti_persona
")
contr.cli <- qimpala(query)
names(contr.cli) <- c("global", "tipo_persona", "ncontr", "ncli")

```

```{r}
contr <- merge(contr, seg, by.x="global", by.y="cod_global")
contr$des_global <- factor(contr$des_global,
                           levels = contr$des_global[order(contr$global)], ordered=TRUE)

contr.cli <- merge(contr.cli, seg, by.x="global", by.y="cod_global")
contr.cli$des_global <- factor(contr.cli$des_global,
                           levels = contr.cli$des_global[order(contr.cli$global)], ordered=TRUE)

```

In total, there are **`r sum(contr.cli$ncli)` customers** that have **`r sum(contr.cli$ncontr)` contracts**. From them, **`r sum(subset(contr.cli, global %in% c(35, 36))$ncli)`** are SMEs with `r sum(subset(contr.cli, global %in% c(35, 36))$ncontr)` contracts.

```{r totalContr}
# by contrapartida
data.table(contr)[,list(ncontr=sum(ncontr), ncli=sum(ncli)), by=c("cod", "contrapartida")][order(cod)]
# pymes by contrapartida
data.table(contr)[global%in%c(35,36),list(ncontr=sum(ncontr), ncli=sum(ncli)), by=c("cod", "contrapartida")][order(cod)]
# by global segment
data.table(contr.cli)[,list(ncontr=sum(ncontr), ncli=sum(ncli)), by=c("global", "des_global")][order(global)]
# by person type
data.table(contr.cli)[,list(ncontr=sum(ncontr), ncli=sum(ncli)), by=c("tipo_persona")]
```

We split the total contracts by Type of Person (natural or legal) and the Global Segment inside the bank.

```{r, fig.height=8, fig.width=10}

contr.tipo <- data.table(contr)[,list(ncontr=sum(ncontr), ncli=sum(ncli)), by=c("tipo_persona", "cod", "contrapartida")]

ggplot(melt(contr.tipo, id=c("tipo_persona", "cod", "contrapartida")), 
       aes(x=contrapartida, y=value, fill=tipo_persona, order=tipo_persona)) + 
  geom_bar(stat="identity", position="dodge") + 
#  scale_y_log10()+
  facet_wrap(~variable) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

#escala logaritmica
ggplot(melt(contr.tipo, id=c("tipo_persona", "cod", "contrapartida")), 
       aes(x=contrapartida, y=value, fill=tipo_persona, order=tipo_persona)) + 
  geom_bar(stat="identity", position="dodge") + 
  scale_y_log10()+
  facet_wrap(~variable) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

# porcentaje
ggplot(melt(contr.tipo, id=c("tipo_persona", "cod", "contrapartida")), 
       aes(x=contrapartida, y=value, fill=tipo_persona, order=tipo_persona)) + 
  geom_bar(stat="identity", position="fill") + 
  facet_wrap(~variable) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

# por segmento global
ggplot(contr, aes(x=contrapartida, y=ncontr, fill=des_global, order=desc(global))) + 
  geom_bar(stat="identity", position="stack") + 
  scale_y_log10()+
  facet_wrap(~tipo_persona, scales="free_y", ncol=1) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

```

We can see that most of those contracts are held mostly by legal customers. The exception is a 25% of pyshical people owning a Operaciones Documentarias contract. 

Now we plot the number of contracts and people only for the SMEs global segments: 35 and 36. 

```{r, fig.height=8, fig.width=12, echo=FALSE}

ggplot(melt(data.table(contr)[global%in%c("35","36"),list(ncontr=sum(ncontr), ncli=sum(ncli)),by=c("des_global", "contrapartida")], 
            id=c("des_global", "contrapartida")), 
       aes(x=contrapartida, y=value, fill=des_global)) + 
  geom_bar(stat="identity") + 
  #scale_y_log10()+
  facet_wrap(~variable, scales="free_y") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
```

### Products

```{r comexClientsProdGest, eval=FALSE}
do.hive("drop table da_pymes.clientes_comex_prod_gest")

q <- paste0("create table da_pymes.clientes_comex_prod_gest as
select cod_entalcli, cod_persctpn, cod_prodgest, cod_std, cod_prtda, descripcion, qnu_contrato
from da_pymes.prod_gest_comex as prod 
  join da_pro.agregacion_comercial_vinculacion agg on trim(cod_prodgest)= cod_prodct
where partition_id= '", month, "' and trim(cod_tipotit)='T' and trim(cod_situdw) in ('A', '')
")
do.hive(q)
do.impala("invalidate metadata da_pymes.clientes_comex_prod_gest")

```



```{r getSegmentosGestion}
query <- paste0("select cod_segmsubo, xti_persona, descripcion, 
  sum(qnu_contrato) ncontr, count(distinct cli.cod_persctpn) ncli
from da_pymes.clientes_comex_prod_gest cli 
  join da_pro.segmento_global sg on sg.cod_persctpn = cli.cod_persctpn and sg.partition_id = '", month, "' and sg.cod_entalfa = '0182'
  join da_pro.clientes_corp corp on corp.cod_persctpn = cli.cod_persctpn and corp.partition_id = '", month, "' and corp.cod_entalfa = '0182'
group by cod_segmsubo, xti_persona, descripcion
")
prod <- qimpala(query)
names(prod) <- c("global", "tipo_persona", "prod_gestion", "ncontr", "ncli")


query <- paste0("select cod_segmsubo, xti_persona,
  count(*) ncontr, count(distinct cli.cod_persctpn) ncli
from da_pymes.clientes_comex_prod_gest cli 
  join da_pro.segmento_global sg on sg.cod_persctpn = cli.cod_persctpn and sg.partition_id = '", month, "' and sg.cod_entalfa = '0182'
  join da_pro.clientes_corp corp on corp.cod_persctpn = cli.cod_persctpn and corp.partition_id = '", month, "' and corp.cod_entalfa = '0182'
group by cod_segmsubo, xti_persona
")
prod.cli <- qimpala(query)
names(prod.cli) <- c("global", "tipo_persona", "ncontr", "ncli")


```


```{r}
prod <- merge(prod, seg, by.x="global", by.y="cod_global")
prod$des_global <- factor(prod$des_global,
                           levels = prod$des_global[order(prod$global)], ordered=TRUE)

prod.cli <- merge(prod.cli, seg, by.x="global", by.y="cod_global")
prod.cli$des_global <- factor(prod.cli$des_global,
                           levels = prod.cli$des_global[order(prod.cli$global)], ordered=TRUE)

```

In total, there are **`r sum(prod.cli$ncli)` customers** that have **`r sum(prod.cli$ncontr)` contracts**. From them, **`r sum(subset(prod.cli, global %in% c(35, 36))$ncli)`** are SMEs with `r sum(subset(prod.cli, global %in% c(35, 36))$ncontr)` contracts.


```{r totalProd}
# by producto
data.table(prod)[,list(ncontr=sum(ncontr), ncli=sum(ncli)), by=c("prod_gestion")][order(prod_gestion)]
# pymes by producto
data.table(prod)[global%in%c(35,36),list(ncontr=sum(ncontr), ncli=sum(ncli)), by=c("prod_gestion")][order(prod_gestion)]
# by global segment
data.table(prod.cli)[,list(ncontr=sum(ncontr), ncli=sum(ncli)), by=c("global", "des_global")][order(global)]
# by person type
data.table(prod.cli)[,list(ncontr=sum(ncontr), ncli=sum(ncli)), by=c("tipo_persona")]
```

We split the total contracts by Type of Person (natural or legal) and the Global Segment inside the bank.

```{r, fig.height=8, fig.width=10}
prod.tipo <- data.table(prod)[,list(ncontr=sum(ncontr), ncli=sum(ncli)), by=c("tipo_persona", "prod_gestion")]

ggplot(melt(prod.tipo, id=c("tipo_persona", "prod_gestion")), 
       aes(x=prod_gestion, y=value, fill=tipo_persona, order=tipo_persona)) + 
  geom_bar(stat="identity", position="stack") + 
#  scale_y_log10()+
  facet_wrap(~variable) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

ggplot(melt(prod, id=c("global", "des_global", "tipo_persona", "prod_gestion")),
       aes(x=prod_gestion, y=value, fill=tipo_persona, order=tipo_persona)) + 
  geom_bar(stat="identity", position="fill") + 
  facet_wrap(~variable) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

ggplot(prod, aes(x=prod_gestion, y=ncontr, fill=des_global, order=desc(global))) + 
  geom_bar(stat="identity", position="stack") + 
  #scale_y_log10()+
  facet_wrap(~tipo_persona, scales="free_y", ncol=1) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

```

```{r onlyPymes, fig.height=8, fig.width=10}

ggplot(subset(prod, global %in% c(35, 36)), aes(x=prod_gestion, y=ncontr, fill=des_global, order=desc(global))) + 
  geom_bar(stat="identity", position="stack") + 
  #scale_y_log10()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

```


We can see that Cheques and Ordenes de pago are used by personal customers mostly, whereas Avales and Remesas by legal ones.


## Let's merge previous approaches

Now we merge together the contracts by *contrapartida* and *producto de gestión*.

```{rpa totalComexClients}
q <- paste0("select cli.*, xti_persona, cod_segmsubo
    from (select cod_persctpn, des_contrapartida as descripcion, 'contrapartida' as type
      from da_pymes.clientes_comex_contrapartidas
      where cod_entalfa = '0182' union all
      select cod_persctpn, descripcion as descripcion, 'producto' as type
      from da_pymes.clientes_comex_prod_gest where cod_entalcli = '0182') cli
  join da_pro.clientes_corp corp on cli.cod_persctpn = corp.cod_persctpn and corp.partition_id = '", month, "' and corp.cod_entalfa = '0182'
  join da_pro.segmento_global sg on sg.cod_persctpn = cli.cod_persctpn and sg.partition_id = '", month, "' and sg.cod_entalfa = '0182'
  ")

cli <- qimpala(q)
```

Some statistics:

```{r}
# Unique customers
length(unique(cli$cod_persctpn))

# SMEs, customers and contracts
length(unique(subset(cli, cod_segmsubo %in% c(35, 36))$cod_persctpn))
nrow(subset(cli, cod_segmsubo %in% c(35, 36)))


cli.n <- data.table(cli)[, list(n=.N, ntype=length(unique(type))), by=c("cod_persctpn")]
summary(data.frame(cli.n)[,c("n", "ntype")])
```

On average, they have `r mean(cli.n$n)` contracts by customer, and `r percent(nrow(subset(cli.n, ntype==2))/nrow(cli.n))` have products by *contrapartida* and *producto de gestión*. 
```{r getComexAgg}

comex.agg <- data.table(cli)[,list(ncontr=.N, npers=length(unique(cod_persctpn))), by=c("descripcion", "type")]
comex.agg.pymes <- data.table(cli)[cod_segmsubo %in% c(35,36),list(ncontr=.N, npers=length(unique(cod_persctpn))), by=c("descripcion", "type")]

comex.agg$descripcion <- factor(comex.agg$descripcion, levels=unique(comex.agg$descripcion)[order(as.character(unique(comex.agg$descripcion)))])
comex.agg.pymes$descripcion <- factor(comex.agg.pymes$descripcion, levels=unique(comex.agg.pymes$descripcion)[order(as.character(unique(comex.agg.pymes$descripcion)))])
  

comex.agg.melt <- melt(comex.agg, id=c("descripcion", "type"))
```

```{r plotMerge, fig.width=12, fig.height=7, echo=FALSE}
ggplot(comex.agg.melt, aes(x=descripcion, y=value, fill=type)) + 
  geom_bar(stat="identity") +  
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + #scale_y_log10() +
  facet_wrap(~variable)

ggplot(comex.agg.pymes, aes(x=descripcion, y=ncontr, fill=type)) + 
  geom_bar(stat="identity") +
  geom_text(aes(y=ncontr, label=ncontr), size=4, vjust=-1) +  
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

ggplot(comex.agg.pymes, aes(x=descripcion, y=ncli, fill=type)) + 
  geom_bar(stat="identity") +
  geom_text(aes(y=ncli, label=ncli), size=4, vjust=-1) +  
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size=12)) 

ggplot(comex.agg, aes(x=descripcion, y=ncontr/npers, fill=type)) + 
  geom_bar(stat="identity") +  
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size=12)) 

```


## Conclusion

We have seen that there are around `r length(unique(subset(cli, cod_segmsubo %in% c('35','36'))$cod_persctpn))` SMEs that have International Trade related activity and that not every type is equally owned. 

We have created an attribute, `comercio_exterior` which denotes whether a BBVA customer has Comex activity or not, avoiding whether it is a SME. The attribute can be seen here: http://20.1.71.9:8787/basics/table/comercio_exterior

