---
title: "Pymes_Explore_Renting"
author: "Marta Lamela"
date: "12/1/2016"
output: html_document
---


```{r, echo=FALSE}
# This is the first mandatory section.

title     <- '[Pymes]: Exploration of Renting contracts'
keywords  <- 'pymes,sme,renting'
```

```{r mandatory, echo=FALSE}
suppressMessages(library(DBI))    # This avoids loading messages and warnings showing up
suppressMessages(library(rJava))
options(warn=-1)

source('~/bda_clarity/tools/warehouse_basics.R')
source('~/bda_clarity/tools/methods_connect.R')
source('~/bda_clarity/tools/write.hive.R')
```

```{r libraries, echo=FALSE}
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(plyr))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(reshape))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(googleVis))
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(reshape))
```

```{r environment}
month <- '20151130'
month_timestamp <- "unix_timestamp('2015-11-30 00:00:00')"
```

```{r dependencies, echo=FALSE}
contrapartida <- clarity.use_table(DEPENDENCY_LASTPART_CATALOGO, 'contrapartida')

global <- clarity.use_table(DEPENDENCY_LASTPART_CATALOGO, 'segmentacion_global', 
                            select='cod_sgmento, des_sgmento')

segmento <- clarity.use_table(DEPENDENCY_LASTPART_PRO_TABLE, 'segmento_global')

intervinientes <- clarity.use_table(DEPENDENCY_LASTPART_PRO_TABLE, 'intervinientes_corp')

clientes <- clarity.use_table(DEPENDENCY_LASTPART_PRO_TABLE, 'clientes_corp')

movimientos <- clarity.use_table(DEPENDENCY_FULL_PRO_TABLE, 'movimientos_cuentas_corp')
```


### *Objective:* PYMES Renting contracts 

### How can we see RENTING *inside* the bank?

There are several RENTING products, and different ways are found in the data:

* RENTING en Epígrafes (cod_prtda):
  + **000006750  .TOTAL RENTING**
  + 000006752	. RENTING AUTO
  + 000006754	. RESTO RENTING
  + 000006756	.  RENTING INMOBILIARIO
  + 000006758	.   EDIFICIOS Y OTRAS CONSTR.
  + 000006760	.   FINCAS RÚSTICAS,PARCELAS Y SOLARES
  + 000006762	.  RENTING MOBILIARIO
  + 000006764	.   EQUIPOS ELECTRONICOS
  + 000006766	.   MOBILIARIO
  + 000006768	.   INSTALACIONES
  + 000006770	.   MATERIAL MOVIL
  + 000006772	.   MAQUINAS
  + 000006774	.   SOFTWARE
  + 000006776	.  RESTO RENTING
  + 000006778	.   RESTO RENTING
  + 000006780	. RESTO RENTING

* RENTING en Contrapartida: 
  + **505: RENTING**
  + 506, 507, 508: RENTING (sin uso actualmente)
  + 355: PROPUESTAS DE RENTING

All this analysis is done for **ACTIVE** renting contracts, identified with cod_situdw = 'A' or NULL (Epígrafes) or fec_cancel not informed and fec_vencto not informed or future (Contrapartida). See *Queries* section below.

Also for the holder **1st TITULAR**. It could be changed with cod_tipotit = 'T' (Epígrafes) or trim(cod_ctippe)='TIT' without cod_ordentit = 1.

```{r createContrapartidaRenting, eval=FALSE}
do.hive("drop table da_pymes.contrapartida_renting")

q <- paste0("create table da_pymes.contrapartida_renting as 
  select cod_pgccontr as cod_contrapartida, max(des_ctrpartd) as des_contrapartida
  from da_catalogos.contrapartida 
  where cod_pgccontr in 
    ('0505', '0506', '0507', '0508')
      and partition_id = '", month, "'
  group by cod_pgccontr
")
do.hive(q)
do.impala("invalidate metadata da_pymes.contrapartida_renting")
```

```{r getContrapartidaRenting, eval=FALSE}
q <- "select * from da_pymes.contrapartida_renting"
dd <- qimpala(q)
```

```{r createProdGestRenting, eval=FALSE}
partida <- c(6750,6752,6754,6756,6758,6760,6762,6764,6766,6768,6770,6772,6774,6776,6778,6780)
estado <- rep(33, length(partida))
descripcion <- c("TOTAL RENTING",
"RENTING AUTO",
"RESTO RENTING",
"RENTING INMOBILIARIO",
"EDIFICIOS Y OTRAS CONSTR.",
"FINCAS RÚSTICAS,PARCELAS Y SOLARES",
"RENTING MOBILIARIO",
"EQUIPOS ELECTRONICOS",
"MOBILIARIO",
"INSTALACIONES",
"MATERIAL MOVIL",
"MAQUINAS",
"SOFTWARE",
"RESTO RENTING",
"RESTO RENTING",
"RESTO RENTING")

epigrafe_renting <- data.frame(estado, partida, descripcion)
do.hive("drop table da_pymes.epigrafe_renting")

write.hive(epigrafe_renting, "da_pymes.epigrafe_renting")
do.impala("invalidate metadata da_pymes.epigrafe_renting")

do.hive("drop table da_pymes.prod_gest_renting")

q <- paste0("create table da_pymes.prod_gest_renting as 
  select cast(cod_prodct as int) as cod_prodct, estado as cod_std, partida as cod_prtda, descripcion
  from da_pro.catalogo_productos_gestion_epigrafes_relacion 
    join da_pymes.epigrafe_renting ep on ep.estado = cast(cod_std as int) and ep.partida = cast(cod_prtda as int) 
    and partition_id = '", month, "'
")
do.hive(q)
do.impala("invalidate metadata da_pymes.prod_gest_renting")
```

```{r getProdGestRenting, eval=FALSE}
q <- "select * from da_pymes.prod_gest_renting"
dd <- qimpala(q)
```


## Quick Exploration

```{r segmentoGlobalCatalogo, cache=TRUE, echo=FALSE}
seg <- qimpala(paste0("select distinct cast(trim(cod_sgmento) as int) cod_global,
               des_sgmento as des_global
from da_catalogos.segmentacion_global 
where partition_id = '", month, "'
"))
```

### Contrapartidas

Let's see the volume of this type of contracts.

```{r rentingClients, eval=FALSE}
do.hive("drop table da_pymes.clientes_renting_contrapartidas")
q <- paste0("create table da_pymes.clientes_renting_contrapartidas as
select cod_entalfa, cod_persctpn, ic.cod_idcontra as cod_idcontra, ic.cod_pgccontr as cod_pgccontr, des_contrapartida
from da_pymes.contrapartida_renting as contr 
join (select cod_entalfa, cod_persctpn, cod_idcontra, cod_pgccontr from da_pro.intervinientes_corp
      where partition_id= '", month, "' and trim(cod_ctippe)='TIT' and cod_ordentit=1 -- and cod_entalfa = '0182'
) ic
  on trim(ic.cod_pgccontr)= cod_contrapartida
join (select cod_pgccontr,cod_idcontra, fec_altacto, fec_vencto, fec_cancel
      from da_pro.productos_contratados where partition_id= '", month, "'  -- and cod_entalfa = '0182'
) pc
  on trim(pc.cod_idcontra) = trim(ic.cod_idcontra)
where fec_cancel is NULL and (fec_vencto is NULL or fec_vencto >= ", month_timestamp, ")
")
do.hive(q)
do.impala("invalidate metadata da_pymes.clientes_renting_contrapartidas")
```

We obtain how many clients and contracts there are for each contrapartida, depending on the customer being Legal or Natural and their Global Segment.

```{r getContrapartidas}
query <- paste0("select cod_segmsubo, xti_persona, cod_pgccontr, cc.des_contrapartida, 
  count(*) ncontr, count(distinct cli.cod_persctpn) ncli
from da_pymes.clientes_renting_contrapartidas cli 
  join da_pymes.contrapartida_renting cc on cc.cod_contrapartida = cli.cod_pgccontr
  join da_pro.segmento_global sg on sg.cod_persctpn = cli.cod_persctpn and sg.partition_id = '", month, "' and sg.cod_entalfa = '0182'
  join da_pro.clientes_corp corp on corp.cod_persctpn = cli.cod_persctpn and corp.partition_id = '", month, "' and corp.cod_entalfa = '0182'
group by cod_segmsubo, xti_persona, 
  cod_pgccontr, cc.des_contrapartida
")
contr <- qimpala(query)
names(contr) <- c("global", "tipo_persona", "cod", "contrapartida", "ncontr", "ncli")
```

```{r}
contr <- merge(contr, seg, by.x="global", by.y="cod_global")
contr$des_global <- factor(contr$des_global,
                           levels = contr$des_global[order(contr$global)], ordered=TRUE)
```

We split the total contracts by Type of Person (Legal or Natural) and the Global Segment inside the bank.

```{r, fig.height=8, fig.width=10}

ggplot(melt(data.table(contr)[,list(ncontr=sum(ncontr), ncli=sum(ncli)),by=c("contrapartida","tipo_persona")],
            id=c("tipo_persona", "contrapartida")), 
       aes(x=contrapartida, y=value, fill=tipo_persona, order=tipo_persona)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~variable) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

ggplot(melt(contr, id=c("global", "des_global", "tipo_persona", "cod", "contrapartida")), 
       aes(x=contrapartida, y=value, fill=tipo_persona, order=tipo_persona)) + 
  geom_bar(stat="identity", position="fill") + 
  facet_wrap(~variable) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

#ggplot(melt(contr, id=c("global", "tipo_persona", "cod", "contrapartida")), 
ggplot(contr, aes(x=contrapartida, y=ncontr, fill=des_global, order=desc(global))) + 
  geom_bar(stat="identity", position="stack") + 
  scale_y_log10()+
  facet_wrap(~tipo_persona, scales="free_y", ncol=2) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

```

We can see that most of those contracts are held mostly by Legal customers.

Now we plot the number of contracts and people only for the SMEs global segments: 35 and 36. 

```{r, fig.height=8, fig.width=12, echo=FALSE}

ggplot(melt(data.table(contr)[global%in%c("35","36"),list(ncontr=sum(ncontr), ncli=sum(ncli)),by=c("des_global", "contrapartida")], 
            id=c("des_global", "contrapartida")), 
       aes(x=contrapartida, y=value, fill=des_global)) + 
  geom_bar(stat="identity") + 
  #scale_y_log10()+
  facet_wrap(~variable, scales="free_y") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
```

Based on this approach, the number of SMEs that have active RENTING related activity is:

```{r resultsContrapartida, echo=FALSE}
data.table(contr)[global%in%c("35","36"),list(ncontr=sum(ncontr), ncli=sum(ncli)),by=c("des_global", "contrapartida")]
```



### Products

```{r rentingClientsProdGest, eval=FALSE}
do.hive("drop table da_pymes.clientes_renting_prod_gest")

q <- paste0("create table da_pymes.clientes_renting_prod_gest as
select cod_entalcli, cod_persctpn, cod_prodgest, cod_std, cod_prtda, descripcion, qnu_contrato
from da_pymes.prod_gest_renting as prod 
  join da_pro.agregacion_comercial_vinculacion agg on trim(cod_prodgest)= cod_prodct
where partition_id= '", month, "' and trim(cod_tipotit)='M' and (trim(cod_situdw)='A' or cod_situdw is NULL)
    -- and cod_entalcli = '0182'
")
do.hive(q)
do.impala("invalidate metadata da_pymes.clientes_renting_prod_gest")
```


```{r getSegmentosGestion}
query <- paste0("select cod_segmsubo, xti_persona, descripcion, 
  sum(qnu_contrato) ncontr, count(distinct cli.cod_persctpn) ncli
from da_pymes.clientes_renting_prod_gest cli 
  join da_pro.segmento_global sg on sg.cod_persctpn = cli.cod_persctpn and sg.partition_id = '", month, "' and sg.cod_entalfa = '0182'
  join da_pro.clientes_corp corp on corp.cod_persctpn = cli.cod_persctpn and corp.partition_id = '", month, "' and corp.cod_entalfa = '0182'
group by cod_segmsubo, xti_persona, descripcion
")
prod <- qimpala(query)
names(prod) <- c("global", "tipo_persona", "prod_gestion", "ncontr", "ncli")
```


```{r}
prod <- merge(prod, seg, by.x="global", by.y="cod_global")
prod$des_global <- factor(prod$des_global,
                           levels = prod$des_global[order(prod$global)], ordered=TRUE)
```


We split the total contracts by Type of Person (Natural or Legal) and the Global Segment inside the bank.

```{r, fig.height=8, fig.width=10}

ggplot(melt(data.table(prod)[,list(ncontr=sum(ncontr), ncli=sum(ncli)),by=c("prod_gestion","tipo_persona")],
            id=c("tipo_persona", "prod_gestion")), 
       aes(x=prod_gestion, y=value, fill=tipo_persona, order=tipo_persona)) + 
  geom_bar(stat="identity", position="stack") + 
  #scale_y_log10()+
  facet_wrap(~variable) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

ggplot(melt(prod, id=c("global", "des_global", "tipo_persona", "prod_gestion")),
       aes(x=prod_gestion, y=value, fill=tipo_persona, order=tipo_persona)) + 
  geom_bar(stat="identity", position="fill") + 
  facet_wrap(~variable) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

ggplot(prod, aes(x=prod_gestion, y=ncontr, fill=des_global, order=desc(global))) + 
  geom_bar(stat="identity", position="stack") + 
  scale_y_log10()+
  facet_wrap(~tipo_persona, scales="free_y", ncol=1) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

```

```{r onlyPymes, fig.height=8, fig.width=10}
ggplot(subset(prod, global%in%c("35","36")), aes(x=prod_gestion, y=ncontr, fill=des_global, order=desc(global))) + 
  geom_bar(stat="identity", position="stack") + 
  #scale_y_log10()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
```

Based on this approach, and considering **TOTAL RENTING** includes all the products related to RENTING, the number of SMEs that have active RENTING activity is:

```{r resultsProdGestion, echo=FALSE}
data.table(prod)[global%in%c("35","36")&prod_gestion%in%c("TOTAL RENTING"),list(ncontr=sum(ncontr), ncli=sum(ncli)),by=c("des_global", "prod_gestion")]
```



## Let's merge previous approaches

Now we merge together the contracts by *contrapartida* and *producto de gestión*.

```{rpa totalrentingClients}
q <- paste0("select cli.*, xti_persona, cod_segmsubo
    from (select cod_persctpn, des_contrapartida as descripcion, 'contrapartida' as type
      from da_pymes.clientes_renting_contrapartidas
      -- where cod_entalfa = '0182'
    union all
      select cod_persctpn, descripcion as descripcion, 'producto' as type
      from da_pymes.clientes_renting_prod_gest where descripcion like '%TOTAL RENTING%'
    -- and cod_entalcli = '0182'
    ) cli
  join da_pro.clientes_corp corp on cli.cod_persctpn = corp.cod_persctpn and corp.partition_id = '", month, "' and corp.cod_entalfa = '0182'
  join da_pro.segmento_global sg on sg.cod_persctpn = cli.cod_persctpn and sg.partition_id = '", month, "' and sg.cod_entalfa = '0182'
  ")

cli <- qimpala(q)
```


There are `r length(unique(cli$cod_persctpn))` clients that have at least one renting product. The average statistics:
```{r}
cli.n <- data.table(cli)[, list(n=.N, ntype=length(unique(type))), by=c("cod_persctpn")]
summary(data.frame(cli.n)[,c("n", "ntype")])
```

On average, **`r percent(nrow(subset(cli.n, ntype==2))/nrow(cli.n))`** of clients have RENTING products by *contrapartida* and *producto de gestión* at the same time.

  + There are `r mean(cli.n$n)` contracts by customer on average, median `r median(cli.n$n)`
  + There are `r mean(subset(cli.n, ntype==1)$n)` in case ntype = 1 (only one class, contrapartida or producto de     gestión informed), `r median(subset(cli.n, ntype==1)$n)`
  + There are `r mean(subset(cli.n, ntype==2)$n)` in case ntype = 2 (two classes informed, contrapartida and producto de gestión), median `r median(subset(cli.n, ntype==2)$n)`

```{r getrentingAgg}

renting.agg <- data.table(cli)[,list(ncontr=.N, npers=length(unique(cod_persctpn))), by=c("descripcion", "type")]

renting.agg.melt <- melt(renting.agg, id=c("descripcion", "type"))
```

```{r plotMerge, fig.width=12, fig.height=7, echo=FALSE}
ggplot(renting.agg.melt, aes(x=descripcion, y=value, fill=type)) + 
  geom_bar(stat="identity") +  
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + #scale_y_log10() +
  facet_wrap(~variable)

ggplot(renting.agg, aes(x=descripcion, y=ncontr/npers, fill=type)) + 
  geom_bar(stat="identity") +  
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

```


And for the SMEs global segments: 35 and 36. 

```{r getrentingAggPymes}
renting.agg.pymes <- data.table(cli)[cod_segmsubo%in%c("35","36"),list(ncontr=.N, npers=length(unique(cod_persctpn))), by=c("descripcion", "type")]

renting.agg.melt.pymes <- melt(renting.agg.pymes, id=c("descripcion", "type"))

cli.n <- data.table(cli)[cod_segmsubo%in%c("35","36"), list(n=.N, ntype=length(unique(type))), by=c("cod_persctpn")]
```

There are `r length(unique(cli$cod_persctpn[cli$cod_segmsubo%in%c("35","36")]))` SMEs that have at least one renting product. On average, **`r percent(nrow(subset(cli.n, ntype==2))/nrow(cli.n))`** of SMEs have RENTING products by *contrapartida* and *producto de gestión* at the same time.


```{r plotMergePymes, fig.width=12, fig.height=7, echo=FALSE}
ggplot(renting.agg.melt.pymes, aes(x=descripcion, y=value, fill=type)) + 
  geom_bar(stat="identity") +  
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + #scale_y_log10() +
  facet_wrap(~variable)

ggplot(renting.agg.pymes, aes(x=descripcion, y=ncontr/npers, fill=type)) + 
  geom_bar(stat="identity") +  
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

```


## Conclusion

We have two options to identify RENTING activity. Based on the data and on the recommendation of MIS, we should use *producto de gestión* because data can be split into types of renting.

For SMEs we have identified renting contracts obtaining similar numbers with both options.

## Queries


```{r COD_ENTALFA, eval=FALSE}

#### ¿USAMOS COD_ENTALFA O COD_ENTALCLI DE AGREGACION_COMERCIAL_VINCULACION? ####

select count(distinct cli.cod_persctpn) ncli
from da_pymes.clientes_renting_prod_gest cli

# 15k clientes con prod gest de RENTING

select count(distinct cli.cod_persctpn) ncli
from da_pymes.clientes_renting_prod_gest cli
join da_pro.segmento_global sg on sg.cod_persctpn = cli.cod_persctpn and sg.partition_id = '20151130' and sg.cod_entalfa = '0182'
join da_pro.clientes_corp corp on corp.cod_persctpn = cli.cod_persctpn and corp.partition_id = '20151130' and corp.cod_entalfa = '0182'

# 15k si hacemos join con segmento global y F/J

select cod_entalcli, cod_entalfa, count(*) n
from da_pro.agregacion_comercial_vinculacion
where partition_id = '20151130'
group by cod_entalcli, cod_entalfa
order by n desc

# hay combinaciones de todo con todo, no tienen por que coincidir los dos códigos

select count(distinct cod_persctpn) ncli
from da_pymes.prod_gest_renting as prod 
  join da_pro.agregacion_comercial_vinculacion agg on trim(cod_prodgest)= cod_prodct
where partition_id= '20151130'
and trim(cod_tipotit)='M'
and cod_entalcli = '0182'

#68k con partition
#21k con partition y entalfa 0182
#55k con partition y menor titular M
#14k con partition y menor titular M y entalcli 0182 ¡PERDEMOS MUCHOS DATOS!

select cod_entalcli, count(distinct cod_persctpn) ncli
from da_pymes.prod_gest_renting as prod 
  join da_pro.agregacion_comercial_vinculacion agg on trim(cod_prodgest)= cod_prodct
where partition_id= '20151130'
and trim(cod_tipotit)='M'
group by cod_entalcli
order by ncli desc

# 0980 42.2k ¡APARECEN NUEVOS CÓDIGOS SIGNIFICATIVOS!
# 0182 14.5k
# 0983 514

select cod_entalfa, count(distinct cod_persctpn) ncli
from da_pymes.prod_gest_renting as prod 
  join da_pro.agregacion_comercial_vinculacion agg on trim(cod_prodgest)= cod_prodct
where partition_id= '20151130'
and trim(cod_tipotit)='M'
group by cod_entalfa
order by ncli desc

# 0980 42.2k
# 0182 14.5k
# 0983 514

select distinct des_nomabr from da_catalogos.entidad
where partition_id = '20151130'
and cod_entalfa = '0980' or cod_entalfa = '0983' 

# BBVA RENTING / AUTORENTING


#### ¿Y QUÉ PASA CON LA CONTRAPARTIDA? ¿OCURRE LO MISMO? #####

select cod_entalfa, count(distinct cod_persctpn) ncli
from da_pymes.contrapartida_renting as contr 
  join da_pro.intervinientes_corp ic on trim(cod_pgccontr)= cod_contrapartida
where partition_id= '20151130'
and trim(cod_ctippe)='TIT' and cod_ordentit=1
group by cod_entalfa
order by ncli desc

# 0980 91k
# 0182 51.2k
# 0983 6.3k
# 0009 122
# 0981 31

select distinct des_nomabr from da_catalogos.entidad
where partition_id = '20151130'
and cod_entalfa = '0980' or cod_entalfa = '0983'
or cod_entalfa = '0009' or cod_entalfa = '0981'

# BBVA RENTING / AUTORENTING / FINANZIA RENTING / FINANZIA BCO CREDITO

```


```{r COD_SITUDW, eval=FALSE}

#### ¿CÓMO CALCULAMOS QUE EL CONTRATO DE RENTING SIGA ACTIVO? #####

select distinct fec_baja
from da_pymes.contrapartida_renting as contr 
  join da_pro.intervinientes_corp ic on trim(cod_pgccontr)= cod_contrapartida
where partition_id= '20151130' and trim(cod_ctippe)='TIT' and cod_ordentit=1
order by fec_baja

# fec_fin_vin y fec_baja a NULL

select distinct fec_vencto
from da_pymes.contrapartida_renting as contr
  join (select cod_pgccontr, cod_idcontra from da_pro.intervinientes_corp where partition_id= '20151130' and trim(cod_ctippe)='TIT' and cod_ordentit=1) ic
  on trim(ic.cod_pgccontr)= cod_contrapartida
  join (select cod_pgccontr,cod_idcontra, fec_vencto from da_pro.productos_contratados where partition_id= '20151130') pc
  on trim(pc.cod_idcontra) = trim(ic.cod_idcontra)

# informado

select distinct fec_altacto, fec_vencto, fec_cancel
from da_pymes.contrapartida_renting as contr
  join (select cod_pgccontr, cod_idcontra from da_pro.intervinientes_corp where partition_id= '20151130' and trim(cod_ctippe)='TIT' and cod_ordentit=1) ic
  on trim(ic.cod_pgccontr)= cod_contrapartida
  join (select cod_pgccontr,cod_idcontra, fec_altacto, fec_vencto, fec_cancel from da_pro.productos_contratados where partition_id= '20151130') pc
  on trim(pc.cod_idcontra) = trim(ic.cod_idcontra)
order by fec_altacto desc, fec_vencto, fec_cancel

# combinaciones para identificar Activos/Cancelados/Vencidos

select count(distinct cod_persctpn) as ncli
from da_pymes.prod_gest_renting as prod 
  join da_pro.agregacion_comercial_vinculacion agg on trim(cod_prodgest)= cod_prodct
where partition_id= '20151130' and trim(cod_tipotit)='M'

# 55735 menor titular

select count(distinct cod_persctpn) as ncli
from da_pymes.prod_gest_renting as prod 
  join da_pro.agregacion_comercial_vinculacion agg on trim(cod_prodgest)= cod_prodct
where partition_id= '20151130' and trim(cod_tipotit)='M' and cod_situdw='A'

# 32383 menor titular activo
# 34796 cualquier titular T
# 46551 cualquiera menos cancelado C
# 32698 cualquiera menos cancelado C y moroso M

select count(distinct cod_persctpn) as ncli
from da_pymes.contrapartida_renting as contr 
  join da_pro.intervinientes_corp ic on trim(cod_pgccontr)= cod_contrapartida
where partition_id= '20151130' and trim(cod_ctippe)='TIT' and cod_ordentit=1

# 137408

select count(distinct cod_persctpn) as ncli
from da_pymes.contrapartida_renting as contr 
join (select cod_pgccontr, cod_persctpn, cod_idcontra from da_pro.intervinientes_corp
      where partition_id= '20151130' and trim(cod_ctippe)='TIT' and cod_ordentit=1) ic
  on trim(ic.cod_pgccontr)= cod_contrapartida
join (select cod_pgccontr,cod_idcontra, fec_altacto, fec_vencto, fec_cancel
      from da_pro.productos_contratados where partition_id= '20151130') pc
  on trim(pc.cod_idcontra) = trim(ic.cod_idcontra)
where fec_cancel is NULL and (fec_vencto is NULL or fec_vencto >= unix_timestamp('2015-11-30 00:00:00'))

# 34509
# 36973 cualquier titular TIT

```
