---
output:
  html_document:
    self_contained: no
    pdf_document: default
    theme: cerulean
---

Digital Sales: Personal Loan
-------------------

```{r title, echo=FALSE}
# This is the first mandatory section.
title     <- '[Digital Sales]: Personal Loan Purpose (Simulator).'

# Keywords. Text separated by commas. E.g. keywords <- 'customer health, recibos' 
# Keywords can classify the attribute in categories such as 'valor, vinculación, digitalidad', describe used tables such as 'recibos, 
# transacciones tarjetas, segmento plan uno' or include other relevant info. Search keywords in the portal to find commonly used terms
# and create new keywords only if justified.

keywords  <- 'loan, omniture, simulator, funnel'
```

```{r loadingPackages, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
fec_inicial <- "2015-06-01"
fec_final   <- "2015-06-30"
fec_datos   <- "2014-06-30'"
suppressMessages(library(DBI))    # This avoids loading messages and warnings showing up
suppressMessages(library(rJava))
options(warn=-1)
source('~/bda_clarity/tools/warehouse_basics.R')
source('~/bda_clarity/tools/write.hive.R')
suppressPackageStartupMessages(library('data.table'))
suppressPackageStartupMessages(library('reshape2'))
suppressPackageStartupMessages(library('ggplot2'))
suppressPackageStartupMessages(library('plyr'))
suppressPackageStartupMessages(library('data.table'))
suppressPackageStartupMessages(library('googleVis'))
suppressPackageStartupMessages(library('gdata'))

op <- options(gvis.plot.tag=NULL)
op <- options(gvis.plot.tag="chart")

doit = FALSE

labels_euro <- function(x) {# no rounding
  paste0(format(x, big.mark = ",", decimal.mark = ".", trim = TRUE,
                scientific = FALSE), " €")
} 


multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  require(grid)
  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

```

```{r dpendencies, echo=FALSE , results='hide'}
### LOADING DEPENDENCIES
omniture.omniture_bbvamice       <- clarity.use_table(DEPENDENCY_OTHER_TABLES, 'omniture.omniture_bbvamice', '*', sqname = 'omniture_bbvamice')
clarity.detalle_productos_saldos_cliente   <- clarity.use_table(DEPENDENCY_CLARITY_TABLE, 'clarity_intermediate.detalle_productos_saldos_cliente', '*', sqname = 'detalle_productos_saldos_cliente')
rms.movimientos_cuentas                    <- clarity.use_table(DEPENDENCY_OTHER_TABLES,    'rms_master.movimientos_cuentas',    '*', sqname = 'movimientos_cuentas')
mmpp.movimientos_tarjetas                  <- clarity.use_table(DEPENDENCY_OTHER_TABLES,    'mmpp_master.movimientos_tarjetas',  '*', sqname = 'movimientos_tarjetas')

```

#### Analysis of personal loan purpose  

March - April 2015 Data (Omniture and CMC)

The goal of this paper is to analyze the purpose of consumer loans and its influence on the process of granting.

```{r querysOmniture, echo=FALSE , cache=TRUE, eval=FALSE}
drop_table('omniture.omniture_cookie_persona')
do.hive(" CREATE TABLE omniture.omniture_cookie_persona AS 
        SELECT
          cod_visitor_id, 
          cod_persona 
        FROM 
          (
          SELECT 
            cod_visitor_id,
            cod_persona, 
            rank() over (PARTITION BY cod_visitor_id ORDER BY num_page_view DESC) as rank
          FROM 
            ( 
            SELECT
              cod_visitor_id, 
              cod_persona, 
              sum(num_page_views) as num_page_view
            FROM omniture.omniture_bbvamice
              where cod_persona!='' AND cod_persona is not null
            GROUP BY 
              cod_visitor_id, 
              cod_persona 
            ) V
          ) R
        WHERE rank=1" )

sql_omniture <- paste0("SELECT a.*,
                       if( a.str_hour = c.str_hour, c.cod_id_camp, NULL ) as str_campaign, 
                       if( a.cod_persona='' or a.cod_persona is null , b.cod_persona,   a.cod_persona ) as  global_id
                       FROM (
                       SELECT * 
                       FROM omniture.omniture_bbvamice
                       WHERE partition_id between '", gsub("-","",as.Date(fec_inicial)) ,"'"," AND '", gsub("-","",fec_final) ,"'"," 
                       ) a
                       LEFT OUTER JOIN  
                       omniture.omniture_cookie_persona b
                       ON (a.cod_visitor_id = b.cod_visitor_id)
                       LEFT OUTER JOIN 
                       ( 
                       SELECT 
                       str_hour, 
                       visitor_id, 
                       collect_set(cod_id_camp) as cod_id_camp
                       FROM da_omniture.finalidad_consumo_reactivo_201506 
                       WHERE cod_id_camp regexp 'consumo reactivo:'
                       GROUP BY str_hour, 
                       visitor_id
                       ) c
                       ON (a.cod_visitor_id = c.visitor_id)
                       WHERE c.visitor_id is not null" )

drop_table("da_rafa.omniture_consumo_reactivo")
do.hive(paste("CREATE TABLE da_rafa.omniture_consumo_reactivo AS ",sql_omniture))
```

```{r loadingCMC, echo=FALSE , cache=TRUE, eval=TRUE}
#loading CMC data and upload to hive table
# cmc_consumo <- read.csv("/us/xe45922/consumo/datos/cruce\ 04\ ult\ tarea\ propuestas\ únicas.csv", header=T, sep =";")
# cmc_consumo_abril <- read.csv("/us/xe45922/consumo/datos/Resultado\ Formularios\ Consumo_20150514.csv", header=T, sep =";"  )
# 
# cmc_consumo_abril$Fecha.Ini <- cmc_consumo_abril$Fecha.Inicio
# names(cmc_consumo) <- gsub("_",".", names(cmc_consumo))
# 
# cmc_consumo <- rbind(cmc_consumo[ , c("Cliente", "DNI.NIF.CIF.del.Cliente","Prod.Comercial","Acontecimiento", "Sub.Acontecimiento", "Fase.Tarea", "Fecha.Fin", "Fecha.Pri.Gen" , "Fecha.Ini")], 
# cmc_consumo_abril[ , c("Cliente", "DNI.NIF.CIF.del.Cliente", "Prod.Comercial","Acontecimiento", "Sub.Acontecimiento", "Fase.Tarea", "Fecha.Fin", "Fecha.Pri.Gen" , "Fecha.Ini")] ) 
# 
# names(cmc_consumo) <- tolower(names(cmc_consumo))
# saved <- write.hive(cmc_consumo , "da_rafa.propuestas_cmc_consumo") 
cmc <- 'da_cmc.propuestas_consumo_20140801_20150729'
cmc_consumo <- qhive(paste("select * from ", cmc ))
```

```{r preprocessingData , cache=TRUE, echo=FALSE}
consumo_reactivo        <- qhive("SELECT * 
                                 FROM da_rafa.omniture_consumo_reactivo 
                                 WHERE str_campaign is not null 
                                 or des_pages regexp 'formulario:simulador prestamos consumo reactivo' ") 
names(consumo_reactivo) <- gsub("omniture_consumo_reactivo.","",names(consumo_reactivo))

consumo_reactivo <- data.table(consumo_reactivo)
consumo_reactivo <- consumo_reactivo[ , id := seq(1,.N)]
setkey(consumo_reactivo, id)

consumo_reactivo$str_campaign  <- gsub("\"|consumo reactivo:|\\[|\\]", "",consumo_reactivo$str_campaign)
consumo_reactivo_id <- consumo_reactivo[, list( titulares   = as.character( strsplit( strsplit( str_campaign, 'titulares:')[[1]][2] , ",")[[1]][1] ) , 
                                                finalidades = as.character( strsplit( strsplit( str_campaign, 'finalidades:')[[1]][2] , ",")[[1]][1] ) ,
                                                cuotas      = as.numeric( gsub("\\.", "", sub("\\.[0-9][0-9]$","", as.character( strsplit( strsplit(str_campaign, 'cuotas:')[[1]][2] , ",")[[1]][1] )))),
                                                ingresos    = as.numeric( gsub("\\.", "", strsplit( strsplit( str_campaign, 'ingresos:')[[1]][2] , ",")[[1]][1] ) )  ,
                                                gastos      = as.numeric( gsub("\\.", "", strsplit( strsplit( str_campaign, 'gastos:')[[1]][2] , ",")[[1]][1] ) ) ,
                                                importes    = as.numeric( gsub("\\.", "", strsplit( strsplit( str_campaign, 'importes:')[[1]][2] , ",")[[1]][1] ) ) ,
                                                plazos      = as.numeric( strsplit( strsplit( str_campaign, 'plazos:')[[1]][2] , ",")[[1]][1]  ) 
                                                ) , by = id]
consumo_reactivo <- consumo_reactivo[consumo_reactivo_id]

consumo_reactivo$cod_persona <- as.numeric(consumo_reactivo$cod_persona)
consumo_reactivo$str_date    <- as.Date(consumo_reactivo$str_date, format="%B %e %Y")

consumo_reactivo <- consumo_reactivo[ , is_client := ifelse(is.na(global_id), 'no cliente', 'cliente')]
consumo_reactivo <- consumo_reactivo[ , global_id := as.numeric( ifelse(is.na(global_id), cod_visitor_id, global_id) ) ]
consumo_reactivo <- consumo_reactivo[ , prestamo  := plazos*cuotas]
consumo_reactivo <- consumo_reactivo[ , intereses := prestamo - importes]
consumo_reactivo <- consumo_reactivo[ , apertura  := importes*0.023]
consumo_reactivo <- consumo_reactivo[ , comisiones:= ceiling(consumo_reactivo$plazo/12*2)/2*40]
consumo_reactivo <- consumo_reactivo[ , a_pagar   := apertura + comisiones + prestamo]
consumo_reactivo <- consumo_reactivo[ , confirmado:=grepl('formulario:simulador prestamos consumo reactivo:4',des_pages) ]

consumo_reactivo <- unique(consumo_reactivo)
# 
# View(consumo_reactivo[consumo_reactivo$confirmado & !is.na(consumo_reactivo$str_campaign), ][order(global_id)])

consumo_reactivo2  <- consumo_reactivo[!is.na(consumo_reactivo$str_campaign),  ][order(global_id , -confirmado, -str_date, -str_hour)]

#consumo_reactivo_form <- consumo_reactivo[consumo_reactivo$confirmado & !is.na(consumo_reactivo$str_campaign), ]
#clientes_cr <- consumo_reactivo_form[ !duplicated(consumo_reactivo_form$global_id), ]
clientes_cr <- consumo_reactivo2[ !duplicated(consumo_reactivo2$global_id), ]
```` 

```{r loadingCRM, echo=FALSE , cache=TRUE, eval=FALSE}
# loading CRM data

form_crm <-qhive("select  substr(regexp_extract(cast(cod_tipident as string) , '([^;]+)',1),20,30) as finalidad, count(distinct cod_persctpn) as total
 from crm_master.datos_cabecera_de_formulario CRM_da 
where cast(CRM_da.cod_formular as string) in ('CRE','PPC','PCC','PCR') 
and trim(substr(cast(cod_user as string),4,6)) = 'MAR-15'
group by substr(regexp_extract(cast(cod_tipident as string), '([^;]+)',1),20,30)")

dt.form_crm <- data.table(form_crm)[order(-total)][-1]
```

Top purposes are housing refurbishing, car (new o used), furniture and travel.


```{r Graphics_simulacion, echo=FALSE, cache=TRUE, results='asis', tidy=FALSE, fig.width=8, fig.height=8}

finalidades <- c("reforma vivienda" , "coche nuevo","coche usado","muebles","viajes","sanidad","estudios","motocicleta","electrodomesticos","boda")
clientes_cr$finalidades <- factor(clientes_cr$finalidades , levels=finalidades )
consumo_reactivo$finalidades <- factor(consumo_reactivo$finalidades, levels =finalidades)

dt.finalidades            <- data.frame( table(clientes_cr[!is.na(clientes_cr$finalidades), ]$finalidades) )
dt.finalidades_confirmado <- data.frame( table(clientes_cr[!is.na(clientes_cr$finalidades) & clientes_cr$confirmado, ]$finalidades) )
names(dt.finalidades)            <- c("loan purpose", "users simulating")
names(dt.finalidades_confirmado) <- c("loan purpose", "users requesting")

dt.finalidades <- suppressMessages( join( dt.finalidades, dt.finalidades_confirmado) ) 

finalidadPrestamo <- gvisBarChart(dt.finalidades, xvar=names(dt.finalidades)[1], yvar=names(dt.finalidades)[-1], 
                       options=list(isStacked=FALSE,
                                    title="Personal loan simulated by clients",
                                    width=1000,
                                    height=500), chartid='finalidades')
plot(finalidadPrestamo)

# 
# finalidadPrestamo_CRM <- gvisBarChart(dt.form_crm, xvar="finalidad", yvar="total", 
#                           options=list(isStacked=FALSE,
#                                     title="Clientes con formularios registrados en CRM por finalidad del préstamo",
#                                     width=1000,
#                                     height=500),  chartid='finalidadesCRM')
# plot(finalidadPrestamo_CRM)

ggplot(clientes_cr[!is.na(clientes_cr$finalidades), ], 
                         aes(y=importes, x = finalidades,  fill = confirmado)) + 
                    geom_boxplot() +  ylab('Loan amount') +
                    ggtitle('Loan amount simulated/requested by purpose') + 
                    scale_y_continuous(labels = labels_euro, limits = c(0,35000)) + 
                    theme_bw(15) + theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

beneficios <- clientes_cr[!is.na(clientes_cr$finalidades) & clientes_cr$confirmado, list(beneficio = sum(intereses, na.rm=T)), by = finalidades][order(-beneficio)]

names(beneficios) <- c('purposes', 'incomes')
InteresesFinalidad <- gvisColumnChart(beneficios, xvar=names(beneficios)[1], yvar=names(beneficios)[-1], 
                       options=list(isStacked=FALSE,
                                    title="incomes by purpose of personal loan request",
                                    width=1000,
                                    height=500), chartid='intereses')
plot(InteresesFinalidad)

```

Users trend to simulate higher loan amount of money that it is requested regardless of the purpose. 

* New car is the purpose for which higher loan amount is requested followed by housing refurbishing.

Based on the number of loans, amount and timing we can see revenues the bank that will receive for each purpose.

```{r preprocessinCMC, cache=TRUE, echo=FALSE, eval=TRUE}
#cmc_consumo <- qhive("select * from da_rafa.propuestas_cmc_consumo") 
names(cmc_consumo)[names(cmc_consumo)=="cliente"] <-  "cod_persona"

# formatting variables
cmc_consumo$fasetarea <- trim(iconv( cmc_consumo$fasetarea, to='ASCII//TRANSLIT'))
cmc_consumo$fasetarea[ cmc_consumo$fasetarea=="Cancelaci?"] <- "Cancelacion"


#Reorder factor
cmc_consumo$fasetarea <- factor(cmc_consumo$fasetarea, levels=c("Cancelacion","Reestudio","Gen.Propuesta","Sanc.Riesgo","Sanc.Precio","Formalizacion","Puesta en Vigor"))
for (k in grep('fec',names(cmc_consumo))){
  cmc_consumo[ , k] <- as.Date( cmc_consumo[ , k]  , "%d/%m/%Y")  
}

# filtering by date and ordering to deduplicate by cod_persona
# cmc_consumo <- cmc_consumo[cmc_consumo$fechaini>=fec_inicial &  cmc_consumo$fechaini<=fec_final , ]
cmc_consumo <- data.table(cmc_consumo)
cmc_consumo <- cmc_consumo[order(cod_persona, -fasetarea, -fechainicio)] 
clientesCMC <- cmc_consumo[ !duplicated(cmc_consumo$cod_persona), ]
```

```{r join_cmc_omniture, echo=FALSE, cache=TRUE, eval=TRUE}
setkey(clientes_cr, global_id)
setkey(clientesCMC, cod_persona)

clientesCMC_cr <- clientesCMC[clientes_cr]
```

#### Personal Loan funnel

```{r GeneralFunnel, echo=FALSE, cache=TRUE, results='asis', tidy=FALSE, fig.width=10, fig.height=10, eval=TRUE}

funnel_general        <- c( dim(clientesCMC_cr)[1] ,  sum(clientesCMC_cr$confirmado), 
                            sum(!is.na(clientesCMC_cr$fasetarea)), sum(clientesCMC_cr$fasetarea == "Puesta en Vigor", na.rm=T))
names(funnel_general) <- c("Simulation", "Request", "contracting process", "Puesta en Vigor")
funnel_general[1:3]   <- abs(diff(funnel_general))
funnel_general        <- data.frame(funnel_general)
funnel_general        <- cbind(row.names(funnel_general), funnel_general)
names(funnel_general) <- c("Last state", "Clients")

FunnelGeneral <- gvisPieChart( data.frame(funnel_general) ,  
                      options   = list(
                        width   = 700,
                        height  = 400,
                        title   = "Personal Loan Funnel",
                        legend  = "{position: 'left', textStyle: {color: 'blue', fontSize: 12}}",
                        pieHole = 0.5), chartid = "FunnelGeneral")
plot(FunnelGeneral)

# How many of this end up in OCL ? 
tipo_prestamos <- cbind( names(data.frame(sort(table(clientesCMC_cr[!is.na(clientesCMC_cr$fasetarea), ]$prodcomercial), decreasing=TRUE))[,1]),
                         data.frame(sort(table(clientesCMC_cr[!is.na(clientesCMC_cr$fasetarea), ]$prodcomercial), decreasing=TRUE))[,1] )
tipo_prestamos <- data.table(tipo_prestamos)
names(tipo_prestamos) <- c("Loan Type", "Clients")
tipo_prestamos$Clients <- as.numeric(tipo_prestamos$Clients)
tipo_prestamos$Pct.Clients <- round(tipo_prestamos$Clients/sum(tipo_prestamos$Clients),3) 

Table <- gvisTable(data.frame(tipo_prestamos) , 
                   formats=list(Clients="#,###",
                                Pct.Clients='##.#%'),
                   options=list( width=600,
                                 height=300,
                                 title="Propuestas CMC"))
plot(Table)
```

<br>

Approximately half loans end up in OCL although it have been requested through the personal loan request.


```{r Graphics_funnel, echo=FALSE, cache=TRUE, results='asis', tidy=FALSE, fig.width=10, fig.height=10, eval=TRUE}

df.Fases <- suppressMessages(dcast(clientesCMC_cr, finalidades ~ fasetarea))
df.Fases <- df.Fases[!is.na(df.Fases$finalidades), ]

funnelConversion <- gvisBarChart(df.Fases, xvar=names(df.Fases)[1], yvar=names(df.Fases)[c( -1, -ncol(df.Fases) ) ], 
                       options=list(isStacked=TRUE,
                                    title="Lead to contract funnel by personal loan purpose",
                                    width=1000,
                                    height=500))
plot(funnelConversion)

df.prop.Fases <- cbind(df.Fases[1], prop.table( as.matrix(df.Fases[c( -1, -ncol(df.Fases) )]) , margin = 1))

funnelConversionPct <- gvisBarChart(df.prop.Fases, xvar=names(df.prop.Fases)[1], yvar=names(df.prop.Fases)[-1], 
                       options=list(isStacked=TRUE,
                                    hAxis="{format:'###,#%'}",
                                    tooltip = "{text:'percentage'}",
                                    title="Lead to contract funnel by personal loan purpose",
                                    width=1000,
                                    height=500))
plot(funnelConversionPct)
```

Study, health or furniture obtains greater conversion rate from request to sale.
On contrary, new car loans obtain lower conversión rate. 

It's important to notice that canceled or newly created requests could lead to a definitive sale soon. 
It doesn't mean request is over. 

```{r saveClientesData, echo=FALSE, eval=TRUE, cache=TRUE, eval=TRUE}
# variables <- c(13,15,16,18,26,28,29,30,35,41:50,68:81)
# write.hive(clientesCMC_cr[with=FALSE , ,variables] , "da_ventadigital.clientes_cr_cmc_20150430")
saved <-  write.hive(clientesCMC_cr , "da_ventadigital.clientes_cr_cmc_20150630") 
```


```{r join_ClientesCR, echo=FALSE, cache=TRUE, echo=FALSE, eval=TRUE }

variables_clientes_cr_cmc <- names(qhive("select * from da_ventadigital.clientes_cr_cmc_20150630 limit 1"))
variables_clientes_cr_cmc <- gsub("clientes_cr_cmc_20150630.","",variables_clientes_cr_cmc)
#paste0(variables_clientes_cr_cmc[ variables_clientes_cr_cmc!="cod_persona"], collapse=",")

variables_clarity_productos <- names(qhive("select * from clarity_intermediate.detalle_productos_saldos_cliente limit 1"))
variables_clarity_productos <- variables_clarity_productos[1:(length(variables_clarity_productos)-1)]

pscCR <- qhive(paste("SELECT * 
                       FROM 
                        (
                          SELECT 
                            ", paste0("consumo_reactivo.",variables_clientes_cr_cmc[ !(variables_clientes_cr_cmc %in% c('cod_ofialfa',"cod_persona")) ], collapse=",") ,", 
                            ", paste0(variables_clarity_productos, collapse=","), "
                          FROM 
                          (
                            SELECT *
                            FROM da_ventadigital.clientes_cr_cmc_20150630
                            WHERE fasetarea='Puesta en Vigor'
                          ) consumo_reactivo
                          INNER JOIN 
                          ", clarity.detalle_productos_saldos_cliente, "
                          ON ( cast(detalle_productos_saldos_cliente.cod_persona as int) = cast(consumo_reactivo.cod_persona as int) ) 
                        ) consumo_detalle
                       LEFT OUTER JOIN 
                        (
                          SELECT * 
                          FROM ", rms.movimientos_cuentas, "
                          WHERE   raw_timestamp >= '20150601000000'
                        ) movimientos_cuentas
                         ON ( trim(consumo_detalle.cod_idcontra) = trim(movimientos_cuentas.cod_idcontra) ) "))
#                         LEFT OUTER JOIN
#                         (
#                           SELECT * 
#                           FROM ", mmpp.movimientos_tarjetas, "
#                           WHERE raw_timestamp >= '20150601000000'
#                         ) movimientos_tarjetas
#                         ON ( trim(consumo_detalle.cod_pan) = trim(movimientos_tarjetas.cod_pan) )" ))

pscCR <- data.table(pscCR)
pscCR$movimientos_cuentas.fec_movimien <- as.Date(pscCR$movimientos_cuentas.fec_movimien)
names(pscCR) <- gsub("consumo_detalle.|movimientos_cuentas.|movimientos_tarjetas.", "", names(pscCR))

```


```{r Graphics_difference, echo=FALSE, cache=TRUE, results='asis', tidy=FALSE, fig.width=7, fig.height=7, eval=TRUE}
pscAbonos <- pscCR[pscCR$des_concepto=='ABONO POR DISPOSICION DE PRESTAMO/CREDITO' & !is.na(pscCR$des_concepto) ,  ]

pscAbonos <- pscAbonos[ pscAbonos$finalidades != "NA" , 
                   list( imp_prestamo = sum(imp_mov) , 
                         imp_solicitado = max(importes), 
                         diferencia = sum(imp_mov) - max(importes), 
                         finalidades = sample(finalidades,1), 
                         fec_ingreso_prestamo = min(fec_movimien)
                         ), by=cod_persona]

pscCR <- join( pscCR, pscAbonos, by = "cod_persona")
pscCR <- pscCR[order(cod_persona, imp_mov)]

ggplot(  pscAbonos, 
        aes(y=diferencia, x=finalidades, colour=finalidades) ) + 
        geom_point() +  ylab('Loan amount difference') +
        ggtitle('Loan amount requested - granted by purpose') + 
        scale_y_continuous(labels = labels_euro, limits = c(-15000,15000)) + 
        theme_bw(15) + theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

```

In many cases there is a significant difference between the amount of borrowed money and current accounts. 

It seems to be a downward rectification of loan amount requested but in some cases the effect is the opposite

```{r expenses_CR, echo=FALSE, cache=TRUE, echo=FALSE, eval=TRUE }
gastosCR <- pscCR[!is.na(imp_prestamo) , list( imp_prestamo = imp_prestamo[1] , 
                                               imp_solicitado = imp_solicitado[1], 
                                               diferencia = diferencia[1], 
                                               finalidades = finalidades[1], 
                                               gasto = abs(imp_mov[1]/imp_solicitado[1]) > 0.1,
                                               fec_ingreso_prestamo = fec_ingreso_prestamo[1] ,
                                               des_concepto = des_concepto[1], 
                                               des_movimto = des_movimto[1],
                                               imp_mov = imp_mov[1] , 
                                               gasto2  = abs(imp_mov[2]/imp_solicitado[2]) > 0.1 ,
                                               des_concepto2 = des_concepto[2], 
                                               des_movimto2 = des_movimto[2],
                                               imp_mov2 = imp_mov[2] ),
                                          by = cod_persona] 

# pscCR <- pscCR[order(cod_persona, -imp_motransa)]
# 
# gastosCR <- pscCR[ !is.na(imp_prestamo) , list( imp_prestamo = imp_prestamo[1] , 
#                                                  imp_solicitado = imp_solicitado[1], 
#                                                  diferencia = diferencia[1], 
#                                                  finalidades = finalidades[1], 
#                                                  gasto = abs(imp_motransa[1]/imp_solicitado[1]) > 0.1,
#                                                  fec_ingreso_prestamo = fec_ingreso_prestamo[1] ,
#                                                  des_sector = des_sector[1], 
#                                                  des_ramo = des_ramo[1],
#                                                  imp_mov = imp_motransa[1] , 
#                                                  gasto2  = abs(imp_motransa[2]/imp_solicitado[2]) > 0.1 ,
#                                                  des_sector2 = cod_sector[2], 
#                                                  des_ramo2 = des_ramo[2],
#                                                  imp_mov2 = imp_motransa[2] ),
#                                             by = cod_persona] 
```


