Integridad sumidero
========================================================

_**In construction**_

```{r, echo = FALSE}
# This is the first mandatory section.

title     <- "[Digital Sales]: How is the data source Sumidero could be beneficial to provide insights to the bank?"

# Keywords. Text separated by commas. E.g. keywords <- 'customer health, recibos' 
# Keywords can classify the attribute in categories such as 'valor, vinculación, digitalidad', describe used tables such as 'recibos, 
# transacciones tarjetas, segmento plan uno' or include other relevant info. Search keywords in the portal to find commonly used terms
# and create new keywords only if justified.

keywords  <- 'sumidero, buzz, logs, trazas'  
```


```{r, echo=FALSE}
# This is the second mandatory section.

suppressMessages(library(DBI))    # This avoids loading messages and warnings showing up
suppressMessages(library(rJava))
suppressMessages(library(ggplot2))
suppressMessages(library(grid))
suppressMessages(library(hexbin))
suppressMessages(library(lattice))

options(warn=-1, scipen=3, width=350)
source('~/bda_clarity/tools/warehouse_basics.R')

```

** PREGUNTA:** ¿Sería posible usar el dato de Sumidero de forma consistente?

Analizamos Diciembre para comprobar el volumen de datos (logs) agregado por horas

```{r results='hide', message=FALSE, cache = TRUE, warning=FALSE, echo=FALSE}

library('ggplot2')

sub1 <- subq.where('substring(datetime,1,13) as variable','sumidero_master.weblogs','substring(datetime,1,13) BETWEEN "2014-12-01" AND "2015-01-01"','sub1')

diciembre <- subq.group('variable, COUNT(variable) as cuenta_logs',sub1,groupby='variable', sqname='diciembre')
Dec <- qhive(paste("select * from",diciembre))

colnames(Dec) <- c("variable", "cuenta_logs")

Dec$asDate <- strptime(Dec$variable, "%Y-%m-%dT%H")

Dec$dia <- format(Dec$asDate,"%d")
Dec$hora <- format(Dec$asDate,"%H")
#Dec$minuto <- substr(Dec$variable,15,16)


DecAgg <- aggregate(cuenta_logs~ hora + dia, data =Dec,sum)


qplot(x=hora,y=cuenta_logs,data=na.omit(DecAgg),color=I('#132B43'), main= 'Sumidero_total_logs') +
  geom_line()+
  facet_wrap(~dia,ncol=8) 

```

Resultado: se observan logs dispares con una desviación fuera del intervalo de confianza.

A continuación buscamos los datos en BUZZ para hacer el mismo resultado

```{r results='hide', message=FALSE, cache = TRUE, warning=FALSE, echo=FALSE}

sub1Buzz <- subq.where('substring(datetime,1,13) as variable','sumidero_master.weblogs','substring(datetime,1,13) BETWEEN "2014-12-01" AND "2015-01-01" AND LCASE(appname) LIKE "%buzz%"','sub1Buzz')

diciembreBuzz <- subq.group('variable, COUNT(variable) as cuenta_logs',sub1Buzz,groupby='variable', sqname='diciembreBuzz')
DecBuzz <- qhive(paste("select * from",diciembreBuzz))

colnames(DecBuzz) <- c("variable", "cuenta_logs")

DecBuzz$asDate <- strptime(DecBuzz$variable, "%Y-%m-%dT%H")

DecBuzz$dia <- format(DecBuzz$asDate,"%d")
DecBuzz$hora <- format(DecBuzz$asDate,"%H")
#Dec$minuto <- substr(Dec$variable,15,16)


DecBuzzAgg <- aggregate(cuenta_logs~ hora + dia, data =DecBuzz,sum)


qplot(x=hora,y=cuenta_logs,data=na.omit(DecBuzzAgg),color=I('#20B2AA'), main= 'Sumidero_Buzz_logs') +
  geom_line() +
  facet_wrap(~dia,ncol=8)

```

Coming soon...


